---
title: "Elk Sightability Camera Report"
subtitle: "An summary of camera trap data from June 2021 to June 2023 on the Sechelt Peninsula"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  word_document:
always_allow_html: TRUE
---

```{r setup and tests, include=FALSE, warnings=F, message=F}

## README FIRST ##
#Read and run this chunk of code line by line in the R Console (do not press knit) - there are some questions below which you will have to answer and some logic tests to complete. Once you are happy that the conditions have been satisfied, hit 'knit' above. 

# Load packages
list.of.packages <-
  c(
    "leaflet",
    "dplyr",
    "kriging",
    "corrplot",
    "lubridate",
    "kableExtra",
    "tidyverse",
    "ggplot2",
    "gridExtra",
    "activity",
    "overlap",
    "webshot",
    "flextable",
    "PNWColors"
  )

# Check you have them and load them
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages, repos = "http://cran.us.r-project.org")
lapply(list.of.packages, require, character.only = TRUE)

# Load your data [change the files paths to your data locations]
load("input/processed_data/SPCS_standard_input.RData")

# Timezone [Use UTC if your cameras do not correct for daylight saving time, if they do use the timezone where the data was collected]
tz <- "UTC"

# Set the "independence" interval in minutes
independent <- 30

# Set a single categorical variable of interest from station covariates for summary graphs. If you do not have an appropriate category, use "project_id".
category <- "project_id"

# Define a colour palette from the package PNWColors
# Can be viewed at https://github.com/jakelawlor/PNWColors#install-package
pal <- "Bay"

# Set focal species (those you want to include in in-depth analysis) and irrelevant species(those you want to exclude from basic results like sex, age, behaviour)
irr_spe <- c("Humans","Dogs","Cats","Unknown species", "Mice", "Bird species")
foc_spe <- c("Elk", "Deer", "Black bears", "Bobcats", "Coyotes", "Wolves")

# Set species which you want collar/tag data for
mrk_spe <- "Elk"

# Set your season dates if you don't have them already or want to change them for this report

ind.dat <- ind.dat %>%
  mutate(month = month(date_time)) %>%
  mutate(season = if_else(month >= 5 & month <= 10, "Summer", "Winter")) %>%
  select(-month)

seas.eff <- row.lookup %>%
  mutate(month = month(date)) %>%
  mutate(season = if_else(month >= 5 & month <= 10, "Summer", "Winter")) %>%
  group_by(station_id, season) %>%
  summarize(eff = n())

# set the dates you want to filter to in this report
start <- "2021-06-15"
end <- "2022-06-01"

# set minimum number of days a camera must be active to be included in this report
min_eff <- 30

# we need to remove any stations with data problems
stations_to_exclude <- c("SPCS17", "SPCS61", "SPCS04")

sta <- sta %>% filter(!station_id %in% stations_to_exclude)
dat <- dat %>% filter(!station_id %in% stations_to_exclude)
ind.dat <- ind.dat %>% filter(!station_id %in% stations_to_exclude)
row.lookup <- row.lookup %>% filter(!station_id %in% stations_to_exclude)
snow.lookup <- snow.lookup %>% filter(!station_id %in% stations_to_exclude)
# effs will update in the next chunk

```

```{r non-adjustable options, echo=F, include=F}
# filter everything based on start/end dates
dat <- dat %>%
  filter(date_time >= as.Date(start) & date_time < as.Date(end))
ind.dat <- ind.dat %>%
  filter(date_time >= as.Date(start) & date_time < as.Date(end))
row.lookup <- row.lookup %>%
  filter(date >= as.Date(start) & date < as.Date(end))
snow.lookup <- snow.lookup %>%
  filter(date >= as.Date(start) & date < as.Date(end))
# now recalculate effs
snow_eff <- snow.lookup %>%
  group_by(station_id) %>%
  summarize(snow_days = n())
eff <- row.lookup %>%
  group_by(station_id) %>%
  summarize(eff = n()) %>%
  full_join(snow_eff, by="station_id") %>%
  mutate(snow_days = replace_na(snow_days, 0)) %>%
  mutate(days_active = eff + snow_days) %>%
# remove data from cameras active less than 'min_eff' days
  filter(eff >= min_eff)
dat <- dat %>%
  filter(station_id %in% eff$station_id)
ind.dat <- ind.dat %>%
  filter(station_id %in% eff$station_id)
sta <- sta %>%
  filter(station_id %in% eff$station_id)


# Count the number of camera stations
n_sta <- length(unique(eff$station_id))

# Generate colours to display the category levels - R needs them as a factor
sta[,category] <- factor(sta[,category])
col.cat <- pnw_palette(pal, (length(levels(sta[,category])) ))
col.cat <- rev(col.cat)
sta$cols <- col.cat[sta[,category]]

# order eff by station_id
eff <- arrange(eff, by="station_id")

# How big should the figures be
eff.height <- 8
if(length(unique(eff$station_id))>80)
   {
     eff.height <- length(unique(eff$station_id))/10
   }

sp.height <- 5
if(length(unique(dat$species))>20)
   {
     sp.height <- 5+(length(unique(dat$species))/8)
   }

# Create a mode function
getmode <- function(v) {
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]
}

```

# Acknowledgements

# Table of Contents

# Executive Summary

# Introduction

## Project Scope

## Background

## Objectives

# Methods

## Study Area

## Camera Field Study

As of `r format(as.Date(start), "%B %Y")`, cameras were active at `r n_sta` unique locations. 

```{r map, fig.height=8, fig.width=6.5, echo=F, fig.cap= "Camera deployment sites."}

m <- leaflet(options = leafletOptions(attributionControl=FALSE, zoomControl=F)) %>%
  addProviderTiles(providers$Esri.WorldImagery, group="satellite") %>%  # Add satellite data
  
  addCircles(lng = sta$longitude, 
                   lat = sta$latitude,
                   color = "white", radius = 250, 
                   opacity = 1,
                   weight=0.5, fillColor = col.cat[1], 
                   fillOpacity = 1,
                   label = paste(sta$station_id))
 # addLegend("bottomleft", colors = col.cat,  
 #           labels = str_to_sentence(levels(sta[,category]), 
 #                                    locale = "en"),
 #           title = str_to_sentence(category),
 #           labFormat = labelFormat(prefix = "$"), 
 #           opacity = 1)

m

allcams <- sta %>% 
  arrange(station_id) %>%
  select(longitude, latitude) %>% 
  as.matrix()

cam_dist <- raster::pointDistance(allcams, allcams, lonlat=T, allpairs = T)
colnames(cam_dist) <- sort(sta$station_id)

cam_dist <- as_tibble(cam_dist) %>%
  mutate(station_id1 = sort(sta$station_id)) %>%
  pivot_longer(cols = c(-station_id1), names_to = "station_id2", values_to = "dist") %>%
  filter(station_id1 != station_id2) %>%
  group_by(station_id1) %>%
  slice_min(dist)

```

Cameras were spaced a mean of `r round(mean(cam_dist$dist))` m from the next nearest camera (± `r round(sd(cam_dist$dist)/sqrt(n_sta),1)` m standard error; range `r round(min(cam_dist$dist))` to `r round(max(cam_dist$dist))` m). Camera effort totaled `r sum(eff$eff)` days across all cameras with a mean of `r round(mean(eff$eff))` active days per camera (± `r round(sd(eff$eff)/sqrt(n_sta),1)` SE; range `r min(eff$eff)` to `r max(eff$eff)`).


# Results

## Raw Camera Detections

To date, `r nrow(dat)` images have undergone classification by species, sex, age, behaviour, and groups size. In total, `r length(levels(factor(dat$species)))-1` species were detected (excluding unknown species). The following figure summarizes detections by species.

```{r detections, echo=F, fig.height=sp.height, fig.width=6.5, fig.cap= "Total detections of each species and proportions of sites each species was detected at."}
layout(matrix(c(1,1,2), 1, 3, byrow = TRUE))
det.sum.total <- dat %>%
  filter(!is.na(report_names)) %>%
  group_by(report_names) %>%
  summarize(n = n(),
            locations = NA) %>%
  arrange(n)

i <-1
for(i in 1:nrow(det.sum.total))
{
  tmp <- subset(dat, report_names==det.sum.total$report_names[i])
  det.sum.total$locations[i] <- length(unique(tmp$station_id))
}

par(mar = c(5, 16, 1, 1))
barplot(
  det.sum.total$n,
  names.arg = paste0(
    det.sum.total$report_names,
    " (n = ",
    det.sum.total$n,
    ", ",
    round((det.sum.total$locations /
             n_sta) * 100),
    "%)"
  ),
  las = 1,
  cex.names = 1,
  xlab = "Total detections",
  horiz = T,
  xlim = c(0, (ceiling(
    max(det.sum.total$n, na.rm = T) / 5000
  ) * 5000))
)

par(mar=c(5,1,1,1))

barplot(det.sum.total$locations/n_sta*100, las=1, cex.names=1, xlab="Percentage of sites detected", horiz=T, xlim=c(0,100))
abline(v=100, lty=2)


```

**Detection Map**

```{r detection map, fig.width=6.5, fig.height=8, echo=F, fig.cap="A map of stations from which images were recovered. Circle size corresponds to the number of animal detections per unit camera effort."}

sta.det <- dat %>%
  filter(!(report_names %in% c(irr_spe))) %>%
  group_by(station_id) %>%
  summarize(det = n()) %>%
  distinct() %>%
  left_join(eff, by="station_id") %>%
  mutate(cr = det/eff*100)

sta.det <- full_join(sta, sta.det, by = "station_id")

scale <- 2

m_det <- leaflet(options = leafletOptions(attributionControl=FALSE, zoomControl=F)) %>%
  addProviderTiles(providers$Esri.WorldImagery, group="Satellite") %>%  # Add satellite data
  addCircles(lng=sta.det$longitude, lat=sta.det$latitude,
             radius=as.numeric(sta.det$cr)*scale,
             weight=1, fillOpacity = 0.6,
             color=col.cat[1],
              label=paste(sta.det$station_id, 
                         sta.det$det, sep=", ")) %>%
  addCircles(lng=sta.det$longitude, lat=sta.det$latitude,
             radius=1,
             weight=1, fillOpacity = 1,
             color="white")
  
m_det

```

## Independent camera detections

Using an independence threshold of `r independent` minutes, the number of independent capture events is `r nrow(ind.dat)`. The rest of the analyses are conducted with this data. A summary of image events is as follows:

```{r ind detections figure, echo=F, eval=T, fig.width=6.5, fig.height=sp.height, fig.cap="Total independent detection events and total animals counted by species."}

layout(matrix(c(1,1,2), 1, 3, byrow = TRUE))
ind.det.sum.total <- as.data.frame(count(ind.dat, report_names)) %>%
  filter(!is.na(report_names))
ind.det.sum.total <- ind.det.sum.total[order(ind.det.sum.total$n),]
ind.det.sum.total <- left_join(ind.det.sum.total, det.sum.total[,c(1,3)], by="report_names")

i <-1
for(i in 1:nrow(ind.det.sum.total))
{
  tmp <- subset(ind.dat, report_names==ind.det.sum.total$report_names[i])
  ind.det.sum.total$animal_count[i] <- sum(tmp$event_groupsize)
}

par(mar=c(5,16,1,1))
barplot(
  ind.det.sum.total$n,
  names.arg = paste0(
    ind.det.sum.total$report_names,
    " (n = ",
    ind.det.sum.total$n,
    ", ",
    ind.det.sum.total$animal_count,
    ")"
  ),
  las = 1,
  cex.names = 1,
  xlab = "Total independent detections",
  horiz = T,
  xlim = c(0, (ceiling(
    max(ind.det.sum.total$n, na.rm = T) / 500
  ) * 500))
)

par(mar = c(5, 1, 1, 1))
barplot(
  ind.det.sum.total$animal_count,
  las = 1,
  cex.names = 1,
  xlab = "Total animals counted",
  horiz = T,
  xlim = c(0, (ceiling(
    max(ind.det.sum.total$animal_count, na.rm = T) / 500
  ) * 500))
)


```

```{r sex ind, echo=F, include=F}
col.name <- "sex"

plot<-FALSE
if(length(colnames(ind.dat)[colnames(ind.dat)==col.name]>0))
   {
      tmp <- table(ind.dat[,col.name], as.character(ind.dat$report_names))
      tmp <- as.data.frame.matrix(tmp)
      tmp <- tmp[ , !(names(tmp) %in% irr_spe)]
      
      ind.dat[,col.name]<- factor(ind.dat[,col.name]) %>%
        droplevels()
      cols <- pnw_palette(pal, length(levels(ind.dat[,col.name])))
      # Name categories with no data N\A for NOT ASSESSED
row.names(tmp)[row.names(tmp)==""] <- "N/A"
      # make it the last level
tmp <- tmp[c(2:nrow(tmp),1),]
      
      data_percentage <- apply(tmp, 2, function(x){x*100/sum(x,na.rm=T)})
      plot<-TRUE
  }

```

```{r sex ind plot, echo=F, fig.width=6.5, fig.height=sp.height, fig.cap="Sex composition of independent detection events by species."}
if(plot==TRUE)
{
layout(matrix(c(1,1,1,2), 1, 4, byrow = TRUE))
par(mar=c(5,10,1,1))
barplot(data_percentage , border="white",col= cols, ylab="", las=1, xlab="% of events", cex.names=0.9, horiz=2)
par(mar=c(0,0,4,0))
plot.new()
legend("topleft", legend=row.names(tmp), fill=cols, xpd=TRUE, cex=1.1 )
} else { print('Not included') }
```

```{r age ind, echo=F, include=F}
col.name <- "age"

plot<-FALSE
if(length(colnames(ind.dat)[colnames(ind.dat)==col.name]>0))
   {
      tmp <- table(ind.dat[,col.name], as.character(ind.dat$report_names))
      tmp <- as.data.frame.matrix(tmp)
      tmp <- tmp[ , !(names(tmp) %in% irr_spe)]
      
      ind.dat[,col.name]<- factor(ind.dat[,col.name])
      cols <- pnw_palette(pal, length(levels(ind.dat[,col.name])))
      # Name catagories with no data N\A for NOT ASSESSED
      row.names(tmp)[row.names(tmp)==""] <- "N/A"
      # make it the last level
      tmp <- tmp[c(2:nrow(tmp),1),]
      
      data_percentage <- apply(tmp, 2, function(x){x*100/sum(x,na.rm=T)})
      plot<-TRUE
  }

```

```{r age ind plot, echo=F, fig.width=6.5, fig.height=sp.height, fig.cap="Age composition of independent detection events by species."}
if(plot==TRUE)
{
layout(matrix(c(1,1,1,2), 1, 4, byrow = TRUE))
par(mar=c(5,10,1,1))
barplot(data_percentage , border="white",col= cols, ylab="", las=1, xlab="% of events", cex.names=0.9, horiz=2)
par(mar=c(0,0,4,0))
plot.new()
legend("topleft", legend=row.names(tmp), fill=cols, xpd=TRUE, cex=1.1 )
} else { print('Not included') }

```

```{r Behaviour ind, echo=F, include=F}
col.name <- "behaviour"
plot<-FALSE
if(length(colnames(ind.dat)[colnames(ind.dat)==col.name]>0))
   {
      tmp <- table(ind.dat[,col.name][!(ind.dat$report_names %in% irr_spe)], as.character(ind.dat$report_names[!(ind.dat$report_names %in% irr_spe)]))
      tmp <- as.data.frame.matrix(tmp)
      tmp$total <- rowSums(tmp)
      tmp <- tmp[tmp$total != 0,]
      tmp$total <- NULL

      
      ind.dat[,col.name]<- factor(ind.dat[,col.name])
      cols <- pnw_palette(pal, nrow(tmp))
      # Name catagories with no data N\A for NOT ASSESSED
      # row.names(tmp)[row.names(tmp)==""] <- "N/A"
      # make it the last level
      # tmp <- tmp[c(2:nrow(tmp),1),]
      
      data_percentage <- apply(tmp, 2, function(x){x*100/sum(x,na.rm=T)})
      plot<-TRUE
  }
```

```{r behaviour ind plot, echo=F, fig.width=6.5, fig.height=sp.height, fig.cap="Behaviours observed in independent detection events by species."}
if(plot==TRUE)
{
layout(matrix(c(1,1,1,2), 1, 4, byrow = TRUE))
par(mar=c(5,10,1,1))
barplot(data_percentage , border="white",col= cols, ylab="", las=1, xlab="% of events", cex.names=0.9, horiz=2)
par(mar=c(0,0,4,0))
plot.new()
legend("topleft", legend=row.names(tmp), fill=cols, xpd=TRUE, cex=1.1 )
} else { print('Not included') }
```

**Independent detection map**

```{r independent map, fig.width=6.5, fig.height=8, echo=F, fig.cap="Count of independent detection events by station. Circle diameter corresponds to event count per unit camera effort."}

sta.deti <- ind.dat %>%
  filter(!(report_names %in% c(irr_spe))) %>%
  group_by(station_id) %>%
  summarize(det = n(),
            animals = sum(event_groupsize)) %>%
  distinct() %>%
  left_join(eff, by="station_id") %>%
  mutate(cr = det/eff*100)

sta.deti <- full_join(sta, sta.deti, by = "station_id")

scale <- 20

m_deti <- leaflet(options = leafletOptions(attributionControl=FALSE, zoomControl=F)) %>%
  addProviderTiles(providers$Esri.WorldImagery, group="Satellite") %>%  # Add satellite data
   
  addCircles(lng=sta.deti$longitude, lat=sta.deti$latitude,
             radius=as.numeric(sta.deti$cr)*scale,
             weight=1, fillOpacity = 0.6,
             color=col.cat[1],
             label=paste(sta.deti$station_id, 
                         sta.deti$det, sep=", ")) %>%
    addCircles(lng=sta.det$longitude, lat=sta.det$latitude,
             radius=1,
             weight=1, fillOpacity = 1,
             color="white")
m_deti

```

**Site-level species covariance**

This plot shows the co variance between different species at the site level for species with \>5 unique detections. A positive value indicates that the two species commonly occur at the same sites, while a negative value indicates that the two species are rarely present at the same site.

```{r covariance, echo=F, fig.width=6, fig.height=sp.height, eval=T, fig.cap="Covariance of independent detections by species."}
par(mfrow=c(1,1))
tmp <- as.data.frame.matrix(table(ind.dat$station_id, ind.dat$report_names))
tmp <- tmp[colSums(tmp)>5]
M <- cor(tmp)

corrplot(M, method="color", col=rev(pnw_palette(pal, 100, type="continuous")),
         type="upper", order="hclust",
         #addCoef.col = "black", # Add coefficient of correlation
         tl.col="black", tl.srt=45, #Text label color and rotation
         # Combine with significance
         #p.mat = p.mat, sig.level = 0.01, insig = "blank",
         # hide correlation coefficient on the principal diagonal
         diag=FALSE
         )

```

### Focal species data

**Distance sampling**

```{r distance hist, fig.width=6, fig.height=3.5, fig.cap="Distances recorded for focal species.", echo=F}
# Filter to focal species
dist <- dat %>%
  filter(!is.na(distance),
         report_names %in% c(foc_spe))
dist %>%
  ggplot(aes(distance)) +
  geom_histogram(color="white", binwidth = 1) +
  facet_wrap(~report_names, scales = "free") +
  labs(x = "Distance to camera (m)", y = "Frequency")
```

#### Elk

**Group size distribution**

```{r group size, echo=F, eval=T, fig.width=6, fig.height=4, fig.cap="Elk group size distribution by sex."}
ind.dat %>%
  filter(report_names == "Elk") %>%
  group_by(sex) %>%
  summarize(mean_group = mean(group_count),
            sd = sd(group_count),
            n = n(),
            st_err = sd/sqrt(n)) %>%
  ggplot(aes(sex, mean_group)) +
  geom_col() +
  scale_x_discrete(name="") +
  scale_y_continuous(name="Average group size") +
  coord_flip() +
  geom_errorbar(aes(x = sex, 
                    ymin = mean_group - st_err, 
                    ymax = mean_group + st_err))

```

**Detection maps**

```{r elk detection map, fig.width=6, fig.height=8, echo=F, message=FALSE, warning=FALSE, fig.cap="Independent elk detection events by station. Circle diameter corresponds to event count per unit camera effort."}

sta.det.elk <- ind.dat %>%
  filter(report_names=="Elk") %>%
  group_by(station_id) %>%
  summarize(det = n(),
            animals = sum(event_groupsize)) %>%
  distinct() %>%
  left_join(eff, by="station_id") %>%
  mutate(cr = det/eff*100)

sta.det.elk <- inner_join(sta, sta.det.elk, by = "station_id")

scale <- 100

m_det.elk <- leaflet(options = leafletOptions(attributionControl=FALSE, zoomControl=F)) %>%
  addProviderTiles(providers$Esri.WorldImagery, group="Satellite") %>%  # Add satellite data
   
  addCircles(lng=sta.det.elk$longitude,
             lat=sta.det.elk$latitude,
             color=pnw_palette(pal, 1),
             radius=as.numeric(sta.det.elk$cr)*scale,
             weight=1, fillOpacity = 0.7,
             label=paste(sta.det.elk$station_id, 
                         sta.det.elk$det, sep=", ")) %>%
    addCircles(lng=sta.det$longitude, lat=sta.det$latitude,
             radius=1,
             weight=1, fillOpacity = 1,
             color="white")
m_det.elk

# mean(sta.det.elk$cr)
# sd(sta.det.elk$cr)/sqrt(nrow(sta.det.elk))

```

```{r elk age detection map, fig.width=6, fig.height=8, echo=F, message=FALSE, warning=FALSE, fig.cap="Independent elk detection events by age composition by station. Circle diameter corresponds to event count per unit camera effort."}

age.det.elk <- ind.dat %>%
  filter(report_names=="Elk") %>%
  mutate(age = if_else(age =="Juvenile", "Adult + Juvenile", age)) %>%
  group_by(station_id, age) %>%
  summarize(det = n()) %>%
  distinct() %>%
  left_join(eff, by="station_id") %>%
  mutate(cr = det/eff*100)

age.det.elk <- inner_join(sta,age.det.elk, by = "station_id")
cols <- pnw_palette(pal, 2)

scale <- 100

m_age.elk <- leaflet(options = leafletOptions(attributionControl=FALSE, zoomControl=F)) %>%
  addProviderTiles(providers$Esri.WorldImagery, group="Satellite") %>%  # Add satellite data

  addCircles(lng=age.det.elk$longitude[age.det.elk$age=="Adult"],
             lat=age.det.elk$latitude[age.det.elk$age=="Adult"],
             radius=as.numeric(age.det.elk$cr[age.det.elk$age=="Adult"])*scale,
             color=cols[1],
             weight=1, fillOpacity = 0.5,
             label=paste(age.det.elk$station_id, 
                         age.det.elk$det, sep=", ")) %>%

  addCircles(lng=age.det.elk$longitude[age.det.elk$age=="Adult + Juvenile"],
             lat=age.det.elk$latitude[age.det.elk$age=="Adult + Juvenile"],
             radius=as.numeric(age.det.elk$cr
                               [age.det.elk$age=="Adult + Juvenile"])*scale,
             weight=1, fillOpacity = 0.5,
             color=cols[2],
             label=paste(age.det.elk$station_id, 
                         age.det.elk$det, sep=", ")) %>%
  
  addCircles(lng=sta.det$longitude, lat=sta.det$latitude,
             radius=1,
             weight=1, fillOpacity = 1,
             color="white") %>%

  addLegend("bottomleft", colors = c(cols),  
            labels = c("Adults", "Adults + Juveniles"),
            title = "Age category",
            labFormat = labelFormat(prefix = "$"), 
            opacity = 1)

m_age.elk

# nrow(age.det.elk[age.det.elk$age != "Adult",])

```

```{r elk seasonal detection map, fig.width=3, fig.height=4, echo=F, message=FALSE, warning=FALSE, fig.cap="Independent elk detection events by station and season. Circle diameter corresponds to event count per unit camera effort."}
seas.elk <- ind.dat %>%
  filter(report_names=="Elk") %>%
  group_by(station_id, season) %>%
  summarize(det = n()) %>%
  distinct() %>%
  left_join(seas.eff, by=c("station_id", "season")) %>%
  mutate(cr = det/eff*100)

seas.elk <- inner_join(sta,seas.elk, by = "station_id")

scale <- 120

m_summer.elk <- leaflet(options = leafletOptions(attributionControl=FALSE, zoomControl=F)) %>%
  addProviderTiles(providers$Esri.WorldImagery, group="Satellite") %>% 
   
  addCircles(lng=seas.elk$longitude[seas.elk$season=="Summer"],
             lat=seas.elk$latitude[seas.elk$season=="Summer"],
             radius=as.numeric(seas.elk$cr[seas.elk$season=="Summer"])*scale,
             weight=0.5, fillOpacity = 0.5, fillColor=cols[2], color=cols[2],
             label=paste(seas.elk$station_id, 
                         seas.elk$cr, sep=", ")) %>%

      addCircles(lng=sta.det$longitude, lat=sta.det$latitude,
             radius=1,
             weight=1, fillOpacity = 1,
             color="white") %>%
  addLegend("bottomleft", colors = cols[c(1,2)],  
           labels = c("Winter", "Summer"),
           title = "Season",
           labFormat = labelFormat(prefix = "$"), 
           opacity = 0.7)

m_summer.elk

m_winter.elk <- leaflet(options = leafletOptions(attributionControl=FALSE, zoomControl=F)) %>%
  addProviderTiles(providers$Esri.WorldImagery, group="Satellite") %>% 

  addCircles(lng=seas.elk$longitude[seas.elk$season=="Winter"],
             lat=seas.elk$latitude[seas.elk$season=="Winter"],
             radius=as.numeric(seas.elk$cr[seas.elk$season=="Winter"])*scale,
             weight=0.5, fillOpacity = 0.5,
             fillColor=cols[1], color=cols[1],
             label=paste(seas.elk$station_id, 
                         seas.elk$cr, sep=", ")) %>%
  
      addCircles(lng=sta.det$longitude, lat=sta.det$latitude,
             radius=1,
             weight=1, fillOpacity = 1,
             color="white")
  
  m_winter.elk
  
  
seas.elk %>% group_by(season) %>% summarize(mean_cr = mean(cr), SE = sd(cr)/sqrt(n()), n = n())

# look at elevation
elev <- read_csv("~/GIS/Cameras/camera_elevations.txt") %>%
  mutate(Name = str_replace(Name, "SP", "SPCS")) %>%
  filter(!(Name %in% stations_to_exclude)) %>%
  select(station_id = Name,
         elevation = ELEVATION)

elk.elev <- ind.dat %>% 
  filter(report_names=="Elk") %>%
  left_join(elev, by="station_id")

seas.elk <- left_join(seas.elk, elev, by="station_id")

seas.elk %>% group_by(season) %>% summarize(mean_elev = weighted.mean(elevation, cr), min = min(elevation), max = max(elevation))

print(summary(aov(elevation ~ season, data = elk.elev)))

```

*Marked animals*

The cameras detected collared individuals in `r length(ind.dat$collar[ind.dat$collar==T])` events across `r length(unique(ind.dat$station_id[ind.dat$collar==T]))` stations, comprising `r length(unique(mrk_spe))` species.

```{r collars, fig.width=6, fig.height=3, include=FALSE}
collars <- ind.dat %>%
  filter(collar==T) %>%
  group_by(collar_tags, station_id) %>%
  summarize(event_count = n()) 

collars <- left_join(collars, sta[,c(1,3,4)], by = "station_id") %>%
  mutate(collar_tags = factor(collar_tags),
         # jitter lat & long so that points aren't on top of each other
         latitude = jitter(latitude, factor = 0.005),
         longitude = jitter(longitude, factor = 0.005))



col.cat <- pnw_palette(pal, (length(levels(collars$collar_tags))))
col.cat <- rev(col.cat)
collars$color <- col.cat[collars$collar_tags]

m_collars <- leaflet(options = leafletOptions(attributionControl=FALSE,
                                              zoomControl=F)) %>%
  addProviderTiles(providers$Esri.WorldImagery, group="Satellite") %>% 
  #  
  # addCircles(lng=sta.det$longitude, lat=sta.det$latitude,
  #            radius=800,
  #            weight=0, fillOpacity = 0.2,
  #            color="white") %>%
  addCircles(lng=collars$longitude,
             lat=collars$latitude,
             radius=300,
             weight=0.5, fillOpacity = 0.9, fillColor=collars$color,
             label = collars$collar_tags) %>%

  addLegend("bottomleft", colors = unique(collars$color),  
           labels = unique(collars$collar_tags),
           title = "Collar ID",
           labFormat = labelFormat(prefix = "$"), 
           opacity = 0.7)

m_collars

```

#### Deer

**Detection maps**

```{r deer detection map, fig.width=6, fig.height=8, echo=F, message=FALSE, warning=FALSE, fig.cap="Independent deer detection events by station. Circle diameter corresponds to event count per unit camera effort."}

sta.det.deer <- ind.dat %>%
  filter(report_names=="Deer") %>%
  group_by(station_id) %>%
  summarize(det = n(),
            animals = sum(event_groupsize)) %>%
  distinct() %>%
  left_join(eff, by="station_id") %>%
  mutate(cr = det/eff*100)

sta.det.deer <- inner_join(sta, sta.det.deer, by = "station_id")

scale <- 40

m_det.deer <- leaflet(options = leafletOptions(attributionControl=FALSE, zoomControl=F)) %>%
  addProviderTiles(providers$Esri.WorldImagery, group="Satellite") %>%  # Add satellite data
   
  addCircles(lng=sta.det.deer$longitude, lat=sta.det.deer$latitude,
             radius=as.numeric(sta.det.deer$cr)*scale,
             weight=1, fillOpacity = 0.7,
             color=cols[1],
             label=paste(sta.det.deer$station_id, 
                         sta.det.deer$det, sep=", ")) %>%
    addCircles(lng=sta.det$longitude, lat=sta.det$latitude,
             radius=1,
             weight=1, fillOpacity = 1,
             color="white")
m_det.deer

```

```{r deer age detection map, fig.width=6, fig.height=8, echo=F, message=FALSE, warning=FALSE, fig.cap="Independent deer detection events by age composition by station. Circle diameter corresponds to event count per unit camera effort."}

age.det.deer <- ind.dat %>%
  filter(report_names=="Deer") %>%
  mutate(age = if_else(age =="Juvenile", "Adult + Juvenile", age)) %>%
  group_by(station_id, age) %>%
  summarize(det = n()) %>%
  distinct() %>%
  left_join(eff, by="station_id") %>%
  mutate(cr = det/eff*100)

age.det.deer <- inner_join(sta,age.det.deer, by = "station_id")

scale <- 45

m_age.deer <- leaflet(options = leafletOptions(attributionControl=FALSE, zoomControl=F)) %>%
  addProviderTiles(providers$Esri.WorldImagery, group="Satellite") %>%  # Add satellite data
   
  addCircles(lng=age.det.deer$longitude[age.det.deer$age=="Adult"],
             lat=age.det.deer$latitude[age.det.deer$age=="Adult"],
             radius=as.numeric(age.det.deer$cr[age.det.deer$age=="Adult"])*scale,
             weight=1, fillOpacity = 0.5, color=cols[1],
             label=paste(age.det.deer$station_id, 
                         age.det.deer$det, sep=", ")) %>%

  addCircles(lng=age.det.deer$longitude[age.det.deer$age=="Adult + Juvenile"],
             lat=age.det.deer$latitude[age.det.deer$age=="Adult + Juvenile"],
             radius=as.numeric(age.det.deer$cr[age.det.deer$age=="Adult + Juvenile"])*scale,
             weight=1, fillOpacity = 0.5,
             color=cols[2],
             label=paste(age.det.deer$station_id, 
                         age.det.deer$det, sep=", ")) %>%
    addCircles(lng=sta.det$longitude, lat=sta.det$latitude,
             radius=1,
             weight=1, fillOpacity = 1,
             color="white") %>%
  addLegend("bottomleft", colors = c(cols),  
           labels = c("Adults", "Adults + Juveniles"),
           title = "Age category",
           labFormat = labelFormat(prefix = "$"), 
           opacity = 1)

m_age.deer

```

```{r deer seasonal detection map, fig.width=3, fig.height=4, echo=F, message=FALSE, warning=FALSE, fig.cap="Independent deer detection events by station and season. Circle diameter corresponds to event count per unit camera effort."}

seas.deer <- ind.dat %>%
  filter(report_names=="Deer") %>%
  group_by(station_id, season) %>%
  summarize(det = n()) %>%
  distinct() %>%
  left_join(seas.eff, by=c("station_id", "season")) %>%
  mutate(cr = det/eff*100)

seas.deer <- inner_join(sta,seas.deer, by = "station_id")

scale <- 50

m_summer.deer <- leaflet(options = leafletOptions(attributionControl=FALSE, zoomControl=F)) %>%
  addProviderTiles(providers$Esri.WorldImagery, group="Satellite") %>% 
   
  addCircles(lng=seas.deer$longitude[seas.deer$season=="Summer"],
             lat=seas.deer$latitude[seas.deer$season=="Summer"],
             radius=as.numeric(seas.deer$cr[seas.deer$season=="Summer"])*scale,
             weight=0.5, fillOpacity = 0.5, fillColor=cols[2], color=cols[2],
             label=paste(seas.deer$station_id[seas.deer$season=="Summer"], 
                         seas.deer$cr[seas.deer$season=="Summer"], sep=", ")) %>%

      addCircles(lng=sta.det$longitude, lat=sta.det$latitude,
             radius=1,
             weight=1, fillOpacity = 1,
             color="white") %>%
  addLegend("bottomleft", colors = cols,  
           labels = c("Winter", "Summer"),
           title = "Season",
           labFormat = labelFormat(prefix = "$"), 
           opacity = 0.7)

m_summer.deer

m_winter.deer <- leaflet(options = leafletOptions(attributionControl=FALSE, zoomControl=F)) %>%
  addProviderTiles(providers$Esri.WorldImagery, group="Satellite") %>% 

  addCircles(lng=seas.deer$longitude[seas.deer$season=="Winter"],
             lat=seas.deer$latitude[seas.deer$season=="Winter"],
             radius=as.numeric(seas.deer$cr[seas.deer$season=="Winter"])*scale,
             weight=0.5, fillOpacity = 0.5,
             fillColor=cols[1], color=cols[1],
             label=paste(seas.deer$station_id[seas.deer$season=="Winter"], 
                         seas.deer$cr[seas.deer$season=="Winter"], sep=", ")) %>%
      addCircles(lng=sta.det$longitude, lat=sta.det$latitude,
             radius=1,
             weight=1, fillOpacity = 1,
             color="white")
  
  m_winter.deer

```

#### Bears

```{r bear detection map, fig.width=6, fig.height=8, echo=F, message=FALSE, warning=FALSE, fig.cap="Independent black bear detection events by station. Circle diameter corresponds to event count per unit camera effort."}

sta.det.bear <- ind.dat %>%
  filter(report_names=="Black bears") %>%
  group_by(station_id) %>%
  summarize(det = n()) %>%
  distinct() %>%
  left_join(eff, by="station_id") %>%
  mutate(cr = det/eff*100)

sta.det.bear <- inner_join(sta, sta.det.bear, by = "station_id")

scale <- 200

m_det.bear <- leaflet(options = leafletOptions(attributionControl=FALSE, zoomControl=F)) %>%
  addProviderTiles(providers$Esri.WorldImagery, group="Satellite") %>%  # Add satellite data
    addCircles(lng=sta.det.bear$longitude, lat=sta.det.bear$latitude,
             radius=as.numeric(sta.det.bear$cr)*scale,
             weight=1, fillOpacity = 0.7,
             color=cols[1],
             label=paste(sta.det.bear$station_id, 
                         sta.det.bear$cr, sep=", ")) %>%
    addCircles(lng=sta.det$longitude, lat=sta.det$latitude,
             radius=1,
             weight=1, fillOpacity = 1,
             color="white")
m_det.bear

```

```{r bear age detection map, fig.width=6, fig.height=8, echo=F, message=FALSE, warning=FALSE, fig.cap="Independent black bear detection events by age composition by station. Circle diameter corresponds to event count per unit camera effort."}

age.det.bear <- ind.dat %>%
  filter(report_names=="Black bears") %>%
  mutate(age = if_else(age =="Juvenile", "Adult + Juvenile", age)) %>%
  group_by(station_id, age) %>%
  summarize(det = n()) %>%
  distinct() %>%
  left_join(eff, by="station_id") %>%
  mutate(cr = det/eff*100)

age.det.bear <- inner_join(sta,age.det.bear, by = "station_id")

scale <- 200

m_age.bear <- leaflet(options = leafletOptions(attributionControl=FALSE, zoomControl=F)) %>%
  addProviderTiles(providers$Esri.WorldImagery, group="Satellite") %>%  # Add satellite data
   
  addCircles(lng=age.det.bear$longitude[age.det.bear$age=="Adult"],
             lat=age.det.bear$latitude[age.det.bear$age=="Adult"],
             radius=as.numeric(age.det.bear$cr[age.det.bear$age=="Adult"])*scale,
             weight=0.5, fillOpacity = 0.4, 
             fillColor=cols[1],
             label=paste(age.det.bear$station_id, 
                         age.det.bear$det, sep=", ")) %>%

  addCircles(lng=age.det.bear$longitude[age.det.bear$age=="Adult + Juvenile"],
             lat=age.det.bear$latitude[age.det.bear$age=="Adult + Juvenile"],
             radius=as.numeric(age.det.bear$cr[age.det.bear$age=="Adult + Juvenile"])*scale,
             weight=0.5, fillOpacity = 0.4,
             fillColor=cols[2],
             label=paste(age.det.bear$station_id, 
                         age.det.bear$cr, sep=", ")) %>%
      addCircles(lng=sta.det$longitude, lat=sta.det$latitude,
             radius=1,
             weight=1, fillOpacity = 1,
             color="white") %>%
  addLegend("bottomleft", colors = c(cols),  
           labels = c("Adults", "Adults + Juveniles"),
           title = "Age category",
           labFormat = labelFormat(prefix = "$"), 
           opacity = 1)

m_age.bear

```

#### Capture rates

```{r relative abundance calc, echo=F, warning=F, message=F, eval=T}

det.sum.site <- as.data.frame(table(ind.dat$station_id[ind.dat$report_names %in% foc_spe], ind.dat$report_names[ind.dat$report_names %in% foc_spe]))
colnames(det.sum.site) <- c("station_id","species", "detections")
det.sum.site$individuals <- NA

i <- 1
for(i in 1:nrow(det.sum.site))
{
   tmp <- subset(ind.dat, station_id==as.character(det.sum.site$station_id)[i] &
              species==as.character(det.sum.site$species)[i])
   det.sum.site$individuals[i] <- sum(tmp$group_count, na.rm=T)
}

# Join with the station effort
CR.site <- left_join(det.sum.site,aggregate(eff~station_id, data=eff,  FUN=sum, na.rm=T) )
CR.site$CR.100 <- round((CR.site$individuals/CR.site$eff)*100,3)
# Add station locations
CR.site <- left_join(CR.site, sta[, c("station_id", "latitude", "longitude")])

```

**Monthly Capture Rates**

Across all focal species, variation in monthly capture rates are as follows:

```{r monthly CR, echo=F, eval=T}
# Capture rates through time
foc_spe <- foc_spe[order(foc_spe)]
# Remove any blanks
foc_spe <- foc_spe[foc_spe!=""]

# Now determine capture rates using the row.lookup
# Make a data frame by month and year
mon.dat <- unique(substr(ind.dat$date_time, 1,7))
mon.dat <- data.frame("month"=mon.dat[order(mon.dat)], "effort"= NA)
mon.dat[as.character(foc_spe)] <- NA
i<-1
for(i in 1:nrow(mon.dat))
{
  mon.dat$effort[i] <- nrow(subset(row.lookup, substr(row.lookup$date,1,7)==mon.dat$month[i]))
  mon.dat$season[i] <- ind.dat$season[substr(ind.dat$date_time,1,7)==mon.dat$month[i]][1]
  mon.dat$total.CR[i] <- (nrow(subset(ind.dat, substr(ind.dat$date_time,1,7)==mon.dat$month[i]))/mon.dat$effort[i])*100
}

# add colors for seasons
mon.dat$cols <- if_else(mon.dat$season == "Winter", cols[1], cols[2])

#exclude months with all cameras active for <5 days
mon.dat <- mon.dat %>%
  filter(effort >= (n_sta*5))

for(i in 1:length(foc_spe))
{
  for(j in 1:nrow(mon.dat))
  {
    tmp <- subset(ind.dat, report_names==as.character(foc_spe)[i] & substr(ind.dat$date_time,1,7)==mon.dat$month[j])
    mon.dat[j, as.character(foc_spe[i])] <- (nrow(tmp)/mon.dat$effort[j])*100
  }
}

mon.dat$timestamp <- strptime(paste0(as.character(mon.dat$month),"-15"), "%Y-%m-%d")

# Remove any silly values 
mon.dat <- mon.dat[is.infinite(mon.dat$total.CR)==F,]

```

```{r monthly CR plot, echo=F, eval=T, fig.width=8, fig.height=3, fig.cap="Camera effort (left) and capture rate of all species (right) per month."}

par(mfrow=c(1,2), mar = c(3,4,1,1))

plot(mon.dat$timestamp, mon.dat$effort, ylab="Effort (camera days)", xlab="Date", type="l", las=1, xaxt="n")
points(mon.dat$timestamp, mon.dat$effort, pch=19, col=rgb(0,0,0,0.4))
axis(1, at = as.numeric(strptime(seq(as.Date(paste0(substr(min(row.lookup$date,
                                                           na.rm=T),
                                                       1,7),"-01")),
                                 as.Date(max(row.lookup$date, na.rm=T)),"months"),
                             "%Y-%m-%d")), 
     labels = format.Date(seq(as.Date(paste0(substr(min(row.lookup$date,
                                                           na.rm=T),
                                                       1,7),"-01")),
                                 as.Date(max(row.lookup$date, na.rm=T)),
                              "months"),"%b"),
     cex.axis = 0.8)

# Overall capture rate
plot(mon.dat$timestamp, mon.dat$total.CR, ylab="Events per 100 days", xlab="Date", type="l", las=1, ylim=c(0, max(mon.dat$total.CR)), xaxt = "n")
points(mon.dat$timestamp, mon.dat$total.CR, pch=19, col=rgb(0,0,0,0.4))
axis(1, at = as.numeric(strptime(seq(as.Date(paste0(substr(min(row.lookup$date,
                                                           na.rm=T),
                                                       1,7),"-01")),
                                 as.Date(max(row.lookup$date, na.rm=T)),"months"),
                             "%Y-%m-%d")), 
     labels = format.Date(seq(as.Date(paste0(substr(min(row.lookup$date,
                                                           na.rm=T),
                                                       1,7),"-01")),
                                 as.Date(max(row.lookup$date, na.rm=T)),
                              "months"),"%b"),
     cex.axis = 0.8)

```

*Species-specific temporal trends*

Species level variation in monthly capture rates are as follows:

```{r species CR, eval=FALSE, fig.height=sp.height, fig.width=6, include=FALSE}

col.cat <- pnw_palette(pal, length(foc_spe))

plot(c(min(mon.dat$timestamp, na.rm=T), max(mon.dat$timestamp, na.rm=T)),
     c(0, max(mon.dat[,3:(2+length(foc_spe))])+5), las=1, ylab="Events per 100 days",  
     xlab="", type="n", yaxt="n", xaxt="n")
axis(1, at = as.numeric(strptime(seq(as.Date(paste0(substr(min(row.lookup$date,
                                                           na.rm=T),
                                                       1,7),"-01")),
                                 as.Date(max(row.lookup$date, na.rm=T)),"months"),
                             "%Y-%m-%d")), 
     labels = format.Date(seq(as.Date(paste0(substr(min(row.lookup$date,
                                                           na.rm=T),
                                                       1,7),"-01")),
                                 as.Date(max(row.lookup$date, na.rm=T)),
                              "months"),"%b"),
     cex.axis = 0.8)

legend("topright", as.character(foc_spe), col=col.cat, lwd = 1)

for(i in 1:length(foc_spe))
{
  lines(mon.dat$timestamp, mon.dat[,as.character(foc_spe)[i]],  type="l", lwd=1.5,
        col=col.cat[i])
  points(mon.dat$timestamp, mon.dat[,as.character(foc_spe)[i]], pch=19, cex=0.8,
         col="black")
}


```

```{r elk CR, echo = F, fig.height=4, fig.width=6, fig.cap="Capture rate of independent elk detection events per 100 days by month"}
par(mar = c(3,4,3,1))
plot(c(min(mon.dat$timestamp, na.rm=T), max(mon.dat$timestamp, na.rm=T)),
     c(0, ceiling(max(mon.dat$Elk)+1)), main="Elk", las=1, ylab="Events per 100 days",  
     xlab="", type="n", yaxt="n", xaxt="n")
axis(1, at = as.numeric(strptime(seq(as.Date(paste0(substr(min(row.lookup$date,
                                                           na.rm=T),
                                                       1,7),"-01")),
                                 as.Date(max(row.lookup$date, na.rm=T)),"months"),
                             "%Y-%m-%d")), 
     labels = format.Date(seq(as.Date(paste0(substr(min(row.lookup$date,
                                                           na.rm=T),
                                                       1,7),"-01")),
                                 as.Date(max(row.lookup$date, na.rm=T)),
                              "months"),"%b"),
     cex.axis = 0.8)
axis(2, at = seq(0, ceiling(max(mon.dat$Elk)), 1))
  lines(mon.dat$timestamp, mon.dat$Elk,  type="l", lwd=1.5)
  points(mon.dat$timestamp, mon.dat$Elk, pch=19,  cex=1, col=mon.dat$cols)
  legend("topright", legend = unique(mon.dat$season), col = unique(mon.dat$cols), pch = 19, bty = "n")
  
mon.dat %>% group_by(season) %>% summarize(mean = mean(Elk), SE = sd(Elk)/sqrt(n()))

# check correlation
print(summary(aov(Elk ~ season, data = mon.dat)))
  
```

```{r bears CR, echo = F, fig.height=3, fig.width=4, fig.cap="Capture rate of independent black bear detection events per 100 days by month"}
par(mar = c(3,4,3,1))
plot(c(min(mon.dat$timestamp, na.rm=T), max(mon.dat$timestamp, na.rm=T)),
     c(0, ceiling(max(mon.dat$'Black bears')+1)), main="Black bears", las=1, ylab="Events per 100 days",  
     xlab="", type="n", yaxt="n", xaxt="n")
axis(1, at = as.numeric(strptime(seq(as.Date(paste0(substr(min(row.lookup$date,
                                                           na.rm=T),
                                                       1,7),"-01")),
                                 as.Date(max(row.lookup$date, na.rm=T)),"months"),
                             "%Y-%m-%d")), 
     labels = format.Date(seq(as.Date(paste0(substr(min(row.lookup$date,
                                                           na.rm=T),
                                                       1,7),"-01")),
                                 as.Date(max(row.lookup$date, na.rm=T)),
                              "months"),"%b"),
     cex.axis = 0.8)
axis(2, at = seq(0, ceiling(max(mon.dat$'Black bears')), 1))
  lines(mon.dat$timestamp, mon.dat$'Black bears',  type="l", lwd=1.5)
  points(mon.dat$timestamp, mon.dat$'Black bears', pch=19,  cex=1, col=mon.dat$cols)
  legend("topright", legend = unique(mon.dat$season), col = unique(mon.dat$cols), pch = 19, bty = "n")  
```

```{r Deer CR, echo = F, fig.height=3, fig.width=4, fig.cap="Capture rate of independent deer detection events per 100 days by month"}
par(mar = c(3,4,3,1))
plot(c(min(mon.dat$timestamp, na.rm=T), max(mon.dat$timestamp, na.rm=T)),
     c(0, ceiling(max(mon.dat$Deer)+1)), main="Deer", las=1, ylab="Events per 100 days",  
     xlab="", type="n", yaxt="n", xaxt="n")
axis(1, at = as.numeric(strptime(seq(as.Date(paste0(substr(min(row.lookup$date,
                                                           na.rm=T),
                                                       1,7),"-01")),
                                 as.Date(max(row.lookup$date, na.rm=T)),"months"),
                             "%Y-%m-%d")), 
     labels = format.Date(seq(as.Date(paste0(substr(min(row.lookup$date,
                                                           na.rm=T),
                                                       1,7),"-01")),
                                 as.Date(max(row.lookup$date, na.rm=T)),
                              "months"),"%b"),
     cex.axis = 0.8)
axis(2, at = seq(0, ceiling(max(mon.dat$Deer)), 2))
  lines(mon.dat$timestamp, mon.dat$Deer,  type="l", lwd=1.5)
  points(mon.dat$timestamp, mon.dat$Deer, pch=19,  cex=1, col=mon.dat$cols)
  legend("topright", legend = unique(mon.dat$season), col = unique(mon.dat$cols), pch = 19, bty = "n")  
```

### Other Species

#### Bobcat

```{r bobcat detection map, fig.width=6, fig.height=8, echo=F, message=FALSE, warning=FALSE, fig.cap="Independent bobcat detection events by station. Circle diameter corresponds to event count per unit camera effort."}

sta.det.bobcat <- ind.dat %>%
  filter(report_names=="Bobcats") %>%
  group_by(station_id) %>%
  summarize(det = n()) %>%
  distinct() %>%
  left_join(eff, by="station_id") %>%
  mutate(cr = det/eff*100)

sta.det.bobcat <- inner_join(sta, sta.det.bobcat, by = "station_id")

scale <- 350

m_det.bobcat <- leaflet(options = leafletOptions(attributionControl=FALSE, zoomControl=F)) %>%
  addProviderTiles(providers$Esri.WorldImagery, group="Satellite") %>%  # Add satellite data
    addCircles(lng=sta.det.bobcat$longitude, lat=sta.det.bobcat$latitude,
             radius=as.numeric(sta.det.bobcat$cr)*scale,
             weight=1, fillOpacity = 0.7,
             color=cols[1],
             label=paste(sta.det.bobcat$station_id, 
                         sta.det.bobcat$cr, sep=", ")) %>%
    addCircles(lng=sta.det$longitude, lat=sta.det$latitude,
             radius=1,
             weight=1, fillOpacity = 1,
             color="white")
m_det.bobcat

```

```{r Bobcats CR, echo = F, fig.height=3, fig.width=4, fig.cap="Capture rate of independent bobcat detection events per 100 days by month"}
par(mar = c(3,4,3,1))
plot(c(min(mon.dat$timestamp, na.rm=T), max(mon.dat$timestamp, na.rm=T)),
     c(0, ceiling(max(mon.dat$Bobcats))), main="Bobcats", las=1, ylab="Events per 100 days",  
     xlab="", type="n", yaxt="n", xaxt="n")
axis(1, at = as.numeric(strptime(seq(as.Date(paste0(substr(min(row.lookup$date,
                                                           na.rm=T),
                                                       1,7),"-01")),
                                 as.Date(max(row.lookup$date, na.rm=T)),"months"),
                             "%Y-%m-%d")), 
     labels = format.Date(seq(as.Date(paste0(substr(min(row.lookup$date,
                                                           na.rm=T),
                                                       1,7),"-01")),
                                 as.Date(max(row.lookup$date, na.rm=T)),
                              "months"),"%b"),
     cex.axis = 0.8)
axis(2, at = seq(0, ceiling(max(mon.dat$Bobcats)), 0.1))
  lines(mon.dat$timestamp, mon.dat$Bobcats,  type="l", lwd=1.5)
  points(mon.dat$timestamp, mon.dat$Bobcats, pch=19,  cex=1, col=mon.dat$cols)
  legend("topright", legend = unique(mon.dat$season), col = unique(mon.dat$cols), pch = 19, bty = "n")  
```

#### Coyote

```{r coyote detection map, fig.width=6, fig.height=8, echo=F, message=FALSE, warning=FALSE, fig.cap="Independent coyote detection events by station. Circle diameter corresponds to event count per unit camera effort."}
par(mar = c(3,4,3,1))
sta.det.coyote <- ind.dat %>%
  filter(report_names=="Coyotes") %>%
  group_by(station_id) %>%
  summarize(det = n()) %>%
  distinct() %>%
  left_join(eff, by="station_id") %>%
  mutate(cr = det/eff*100)

sta.det.coyote <- inner_join(sta, sta.det.coyote, by = "station_id")

scale <- 250

m_det.coyote <- leaflet(options = leafletOptions(attributionControl=FALSE, zoomControl=F)) %>%
  addProviderTiles(providers$Esri.WorldImagery, group="Satellite") %>%  # Add satellite data
    addCircles(lng=sta.det.coyote$longitude, lat=sta.det.coyote$latitude,
             radius=as.numeric(sta.det.coyote$cr)*scale,
             weight=1, fillOpacity = 0.7,
             color=cols[1],
             label=paste(sta.det.coyote$station_id, 
                         sta.det.coyote$cr, sep=", ")) %>%
    addCircles(lng=sta.det$longitude, lat=sta.det$latitude,
             radius=1,
             weight=1, fillOpacity = 1,
             color="white")
m_det.coyote

```

```{r Coyotes CR, echo = F, fig.height=3, fig.width=4, fig.cap="Capture rate of independent coyote detection events per 100 days by month"}
par(mar = c(3,4,3,1))

plot(c(min(mon.dat$timestamp, na.rm=T), max(mon.dat$timestamp, na.rm=T)),
     c(0, ceiling(max(mon.dat$Coyotes))), main="Coyotes", las=1, ylab="Events per 100 days",  
     xlab="", type="n", yaxt="n", xaxt="n")
axis(1, at = as.numeric(strptime(seq(as.Date(paste0(substr(min(row.lookup$date,
                                                           na.rm=T),
                                                       1,7),"-01")),
                                 as.Date(max(row.lookup$date, na.rm=T)),"months"),
                             "%Y-%m-%d")), 
     labels = format.Date(seq(as.Date(paste0(substr(min(row.lookup$date,
                                                           na.rm=T),
                                                       1,7),"-01")),
                                 as.Date(max(row.lookup$date, na.rm=T)),
                              "months"),"%b"),
     cex.axis = 0.8)
axis(2, at = seq(0, ceiling(max(mon.dat$Coyotes)), 0.5))
  lines(mon.dat$timestamp, mon.dat$Coyotes,  type="l", lwd=1.5)
  points(mon.dat$timestamp, mon.dat$Coyotes, pch=19,  cex=1, col=mon.dat$cols)
  legend("topright", legend = unique(mon.dat$season), col = unique(mon.dat$cols), pch = 19, bty = "n")  
```

#### Wolf

```{r wolf detection map, echo=FALSE, fig.cap="Independent coyote detection events by station. Circle diameter corresponds to event count per unit camera effort.", fig.height=8, fig.width=6, message=FALSE, warning=FALSE}

sta.det.wolf <- ind.dat %>%
  filter(report_names=="Wolves") %>%
  group_by(station_id) %>%
  summarize(det  = n()) %>%
  distinct() %>%
  left_join(eff, by="station_id") %>%
  mutate(cr = det/eff*100)

sta.det.wolf <- inner_join(sta, sta.det.wolf, by = "station_id")

scale <- 2000

m_det.wolf <- leaflet(options = leafletOptions(attributionControl=FALSE, zoomControl=F)) %>%
  addProviderTiles(providers$Esri.WorldImagery, group="Satellite") %>%  # Add satellite data
    addCircles(lng=sta.det.wolf$longitude, lat=sta.det.wolf$latitude,
             radius=as.numeric(sta.det.wolf$cr)*scale,
             weight=1, fillOpacity = 0.7,
             color=cols[1],
             label=paste(sta.det.wolf$station_id, 
                         sta.det.wolf$cr, sep=", ")) %>%
    addCircles(lng=sta.det$longitude, lat=sta.det$latitude,
             radius=1,
             weight=1, fillOpacity = 1,
             color="white")
m_det.wolf

```

```{r Wolves CR, echo=FALSE, fig.cap="Capture rate of independent coyote detection events per 100 days by month", fig.height=3, fig.width=4}
par(mar = c(3,4,3,1))

plot(c(min(mon.dat$timestamp, na.rm=T), max(mon.dat$timestamp, na.rm=T)),
     c(0, max(mon.dat$Wolves)), main="Wolves", las=1, ylab="Events per 100 days",  
     xlab="", type="n", yaxt="n", xaxt="n")
axis(1, at = as.numeric(strptime(seq(as.Date(paste0(substr(min(row.lookup$date,
                                                           na.rm=T),
                                                       1,7),"-01")),
                                 as.Date(max(row.lookup$date, na.rm=T)),"months"),
                             "%Y-%m-%d")), 
     labels = format.Date(seq(as.Date(paste0(substr(min(row.lookup$date,
                                                           na.rm=T),
                                                       1,7),"-01")),
                                 as.Date(max(row.lookup$date, na.rm=T)),
                              "months"),"%b"),
     cex.axis = 0.8)
axis(2, at = seq(0, max(mon.dat$Wolves), 0.05))
  lines(mon.dat$timestamp, mon.dat$Wolves,  type="l", lwd=1.5)
  points(mon.dat$timestamp, mon.dat$Wolves, pch=19,  cex=1, col=mon.dat$cols)
  legend("topright", legend = unique(mon.dat$season), col = unique(mon.dat$cols), pch = 19, bty = "n")  
```

### Human data

```{r human data, echo=F, message=FALSE, warning=FALSE, include=F}
human.dat <- ind.dat %>%
  filter(species=="Homo sapiens") %>%
  group_by(station_id) %>%
  summarize(events = n(),
            most_popular=as.character(getmode(behaviour)))

human.sta <- human.dat %>%
  full_join(sta, by="station_id") %>%
  mutate(most_popular = if_else(is.na(most_popular), NA, most_popular))

human.sta$most_popular <- factor(human.sta$most_popular)

col.cat <- pnw_palette(pal, length(levels(human.sta$most_popular)))
human.sta$cols <- col.cat[human.sta$most_popular]

```

```{r human map, fig.width=6, fig.height=8, echo=F, fig.cap="Independent human detection events by station. Circle diameter corresponds to event count per unit camera effort. Circle color indicates the most common human use type observed."}

m_hum <- leaflet(options = leafletOptions(attributionControl=FALSE, zoomControl=F)) %>%
  addProviderTiles(providers$Esri.WorldImagery, group="satellite") %>%  # Add satellite data
  
  addCircles(lng = human.sta$longitude, 
                   lat = human.sta$latitude,
                   fillColor = human.sta$cols, 
                   color=human.sta$cols,
                   radius = (if_else(is.na(human.sta$events), 0,
                                     200+human.sta$events*50)),
                   weight=0.5,
                   fillOpacity = 0.5,
                   label = c(human.sta$station_id, human.sta$events)) %>%
  
      addCircles(lng=sta.det$longitude, lat=sta.det$latitude,
             radius=50,
             weight=1, fillOpacity = 1,
             color="white") %>%
  
 addLegend("bottomleft", colors = c(col.cat[seq(1,length(levels(human.sta$most_popular)))]),  
           labels = c(levels(human.sta$most_popular)),
           title = "Most popular activity",
           labFormat = labelFormat(prefix = "$"), 
           opacity = 1)
m_hum

```

# Appendix A

## Species names

```{r species names, echo=F, include=T}
species.names <- dat %>%
  ungroup() %>%
  select(species, common_names, report_names) %>%
  unique() %>%
  filter(common_names != "") %>%
  arrange(species) %>%
  transmute("Latin name" = species, "Common names" = common_names, "Used in report" = report_names)

flextable(species.names, cwidth = c(1.5, 3, 1.5))

```

# Appendix B

## Classification definitions

# Appendix C

## Camera activity

```{r cam activity, echo=F, fig.height=9, fig.width=6.5}

# Adjust layout
par(mar=c(2,6,1,1))
plot(c(min(row.lookup$date, na.rm=T), 
       max(row.lookup$date+1, na.rm=T)), 
       c(1,n_sta+5), las=1, ylab="", xlab="", type="n", yaxt="n", xaxt="n")

# Have the first station plot at the top 
plot.order <- rev(unique(eff$station_id))

axis(2, at= 1:n_sta, labels= plot.order, las=1, cex.axis=0.6)
col.cat <- pnw_palette("Bay", 4)

# Vertical lines for months
abline(v = 
  seq(as.Date(paste0(
    substr(min(row.lookup$date,
               na.rm = T),
           1, 7), "-01"
  )),
  as.Date(max(
    row.lookup$date, na.rm = T
  ))+1, "months"), col = rgb(0, 0, 0, 0.1))

axis(
  1,
  at = seq(as.Date(paste0(
      substr(min(row.lookup$date,
                 na.rm = T),
             1, 7), "-01"
    )),
    as.Date(max(
      row.lookup$date, na.rm = T
    ))+1, "months"),
  labels = format.Date(seq(
    as.Date(paste0(substr(
      min(row.lookup$date,
          na.rm = T),
      1, 7
    ), "-01")),
    as.Date(max(row.lookup$date, na.rm = T))+1,
    "months"
  ), "%b"),
  cex.axis = 0.8
)

mtext("Station ID", 2, 4, cex = 1)
# Make lines for each of the cameras
for(i in 1:length(plot.order))
{
  abline(h=i, col=rgb(0,0,0,0.1))
  tmp <- row.lookup[row.lookup$station_id==plot.order[i],]
  for(j in 1:nrow(tmp))
    {
      lines(c(tmp$date[j],
                       tmp$date[j]+days(1)),
            c(i,i), lwd=2)
  }
}

for(i in 1:length(plot.order))
{
  tmp <- snow.lookup[snow.lookup$station_id == plot.order[i],]
  for (j in 1:nrow(tmp))
  {
    if (nrow(tmp) > 0) {
      lines(c(tmp$date[j],
              tmp$date[j] + days(1)),
            c(i, i),
            lwd = 4,
            col = "dodgerblue2")
    }
  }
}

# legend
legend("topright", inset = 1/50, c("Active", "Active, covered in snow"), lwd = c(2, 4), col = c("black", "dodgerblue2"))

```

# Appendix D

## Detection check

```{r det species colors, include=F}
library(stringr)
# Make species colour codes
species.colors <-
  data.frame("species" = str_sort(unique(dat$report_names[!is.na(dat$report_names)])),
             "Colour" = as.character(pnw_palette(pal, length(
               unique(dat$report_names[!is.na(dat$report_names)])
             ))))

```

```{r detecion summary, echo=F, message=F, warning=F, fig.width=6.5, fig.height=9}

# Make a separate plot for each station
# To do this make a plot dataframe
tmp_plot <-
  data.frame("station_id" = rev(plot.order),
             "Plot.grp" = ceiling(1:length(unique(eff$station_id)) / 20))


tmp_eff_plot <- left_join(eff, tmp_plot, by = "station_id")
i <- 1
j <- 1
for (j in 1:length(unique(tmp_eff_plot$Plot.grp)))
{
  tmp_sta <- rev(unique(tmp_eff_plot$station_id[tmp_eff_plot$Plot.grp == j]))
  k <- length(tmp_sta)
  
  layout(matrix(c(1, 1, 1, 2), 1, 4, byrow = TRUE))
  par(mar = c(2, 6, 1, 1))
  
  plot(
    c(
      min(row.lookup$date, na.rm = T),
      max(row.lookup$date, na.rm = T)
    ),
    c(1, k),
    las = 1,
    ylab = "",
    xlab = "",
    type = "n",
    yaxt = "n"
  )
  # vertical lines for months
  abline(v = 
  seq(as.Date(paste0(
    substr(min(row.lookup$date,
               na.rm = T),
           1, 7), "-01"
  )),
  as.Date(max(
    row.lookup$date, na.rm = T
  )), "months"), col = rgb(0, 0, 0, 0.1))

  axis(
    2,
    at = 1:k,
    labels = tmp_sta,
    las = 1,
    cex.axis = 1
  )
  #mtext("Camera Deployment ID", 2, 4)
  # Make lines for each of the cameras
  for (i in 1:k)
  {
    abline(h = i, col = rgb(0, 0, 0, 0.1))
    tmp_eff <- eff[eff$station_id == tmp_sta[i], ]
    tmp_dat <- dat[dat$station_id == tmp_sta[i], ]
    tmp_dat <-
      left_join(tmp_dat, species.colors, by = c("report_names" = "species")) %>%
      mutate(date_time = as.Date(date_time))
    
    points(tmp_dat$date_time,
           rep(i, nrow(tmp_dat)),
           pch = "|",
           col = tmp_dat$Colour)
    
    tmp_lookup <-
      row.lookup[row.lookup$station_id == tmp_sta[i], ]
    for (m in 1:nrow(tmp_lookup))
    {
      lines(c(tmp_lookup$date[m],
              tmp_lookup$date[m] + days(1)),
            c(i, i), lwd = 2)
    }
  }
  par(mar = c(0, 0, 1, 0))
  plot.new()
  legend(
    "topleft",
    legend = species.colors$species,
    fill = species.colors$Colour,
    xpd = TRUE,
    cex = 1.1
  )
  
}

```

```{r echo=F, eval=F}
### Data exporting for analysis
Finally, this script outputs 10 useful data frames for future data analysis:

1. A data frame of "independent detections" at the `r independent` minute threshold you specified at the start: 

  - "`r  paste0("processed_data/",dat$project_id[1], "_",independent ,"min_Independent.csv")`"
  
2. The "effort lookup" which is a dataframe of all days a given camera station was active. Some people use an effort matrix for this step, but we find the long format is much easier to use in downstream analysis. 
  - "`r paste0("processed_data/",dat$project_id[1], "_daily_effort_lookup.csv")`"  

3 & 4: A 'site x species' matrix of the number of independent detections and species counts across the full study period:

  - "`r  paste0("processed_data/",dat$project_id[1], "_",independent ,"min_Independent_total_observations.csv")`"

  - "`r  paste0("processed_data/",dat$project_id[1], "_",independent ,"min_Independent_total_counts.csv")`"


5 & 6: A 'site_month x species' matrix of the number of independent detections and species counts across for each month in the study period:

  - "`r  paste0("processed_data/",dat$project_id[1], "_",independent ,"min_Monthly_total_observations.csv")`"

  - "`r  paste0("processed_data/",dat$project_id[1], "_",independent ,"min_Monthly_total_counts.csv")`"


7 & 8: A 'site_week x species' matrix of the number of independent detections and species counts across for each week in the study period:

  - "`r  paste0("processed_data/",dat$project_id[1], "_",independent ,"min_Weekly_total_observations.csv")`"

  - "`r  paste0("processed_data/",dat$project_id[1], "_",independent ,"min_Weekly_total_counts.csv")`"
  
9 & 10: A 'site_day x species' matrix of the number of independent detections and species counts across for each day a station was active in the study period:

  - "`r  paste0("processed_data/",dat$project_id[1], "_",independent ,"min_Daily_total_observations.csv")`"

  - "`r  paste0("processed_data/",dat$project_id[1], "_",independent ,"min_Daily_total_counts.csv")`"
  
```

```{r, echo=F, eval=FALSE}
# Only do this for species you are interested in
ind.dat <- ind.dat[ind.dat$species %in% foc_spe,]
ind.dat$species <- factor(ind.dat$species)
#levels(ind.dat$species)
```

```{r, echo=F, message=F, warning=F, eval=F}
# Total counts
  # Station / Month / Effort / Species      
  tmp <- row.lookup
  
  # Calculate the number of days at each site  
  total.obs <- tmp %>% 
      group_by(station_id) %>%
      summarise(effort = n())
  # Convert to a data frame
  total.obs <- as.data.frame(total.obs)
  
  # Add columns for each species  
  total.obs[, levels(ind.dat$species)] <- NA
  # Duplicate for counts
  total.count <- total.obs
  # Test counter
  i <-1
  # For each station, count the number of individuals/observations
  for(i in 1:nrow(total.obs))
    {
      tmp <- ind.dat[ind.dat$station_id==total.obs$station_id[i],]
      
      for(j in 1:length(levels(ind.dat$species)))
      {
        total.obs[i,levels(ind.dat$species)[j]] <- length(tmp$species[tmp$species==levels(ind.dat$species)[j]])
        total.count[i,levels(ind.dat$species)[j]] <- sum(tmp$event_groupsize[tmp$species==levels(ind.dat$species)[j]])
      }
    }

#write.csv(total.obs, paste0("processed_data/",dat$project_id[1], "_",independent ,"min_Independent_total_observations.csv"), row.names = F) 

#write.csv(total.count, paste0("processed_data/",dat$project_id[1], "_",independent ,"min_Independent_total_counts.csv"), row.names = F) 

```

```{r, echo=F, message=F, warning=F, eval=F}

# Monthly counts
  # Station / Month / Effort / Covariates / Species      
  tmp <- row.lookup
  # Simplify the date to monthly
  tmp$date <- substr(tmp$date,1,7)
  
  # Calculate the number of days in each month  
  mon.obs <- tmp %>% 
      group_by(station_id,date ) %>%
      summarise(effort = n())
  # Convert to a data frame
  mon.obs <- as.data.frame(mon.obs)
    
  mon.obs[, levels(ind.dat$species)] <- NA
  mon.count <- mon.obs
  # For each month, count the number of individuals/observations
  for(i in 1:nrow(mon.obs))
    {
      tmp <- ind.dat[ind.dat$station_id==mon.obs$station_id[i] & substr(ind.dat$date_time,1,7)== mon.obs$date[i],]
      for(j in 1:length(levels(ind.dat$species)))
      {
        mon.obs[i,levels(ind.dat$species)[j]] <- length(tmp$species[tmp$species==levels(ind.dat$species)[j]])
        mon.count[i,levels(ind.dat$species)[j]] <- sum(tmp$event_groupsize[tmp$species==levels(ind.dat$species)[j]])
      }
    }

  
#write.csv(mon.obs, paste0("processed_data/",dat$project_id[1], "_",independent ,"min_Independent_Monthly_observations.csv"), row.names = F) 

#write.csv(mon.count, paste0("processed_data/",dat$project_id[1], "_",independent ,"min_Independent_Monthly_counts.csv"), row.names = F) 

```

```{r, echo=F, message=F, warning=F, eval=F}
# Weekly format
  # Station / Month / Effort / Covariates / Species      
  tmp <- row.lookup
  # Simplify the date to year-week
  tmp$date <- strftime(tmp$date, format = "%Y-W%U")
  # The way this is coded is the counter W01 starts at the first sunday of the year, everything before that is W00. Weeks do not roll accross years.
  
  # Calculate the number of days in each week  
  week.obs <- tmp %>% 
      group_by(station_id,date ) %>%
      summarise(effort = n())
  
  # Convert to a data frame
  week.obs <- as.data.frame(week.obs)
  # Add species columns  
  week.obs[, levels(ind.dat$species)] <- NA
  week.count <- week.obs
  # For each week, count the number of individuals/observations
  for(i in 1:nrow(week.obs))
    {
      tmp <- ind.dat[ind.dat$station_id==week.obs$station_id[i] & strftime(ind.dat$date_time, format = "%Y-W%U")== week.obs$date[i],]
      
      
      
      for(j in 1:length(levels(ind.dat$species)))
      {
        week.obs[i,levels(ind.dat$species)[j]] <- length(tmp$species[tmp$species==levels(ind.dat$species)[j]])
        week.count[i,levels(ind.dat$species)[j]] <- sum(tmp$event_groupsize[tmp$species==levels(ind.dat$species)[j]])
      }
    }
#write.csv(week.obs, paste0("processed_data/",dat$project_id[1], "_",independent ,"min_Independent_weekly_observations.csv"), row.names = F) 

#write.csv(week.count, paste0("processed_data/",dat$project_id[1], "_",independent ,"min_Independent_weekly_counts.csv"), row.names = F) 

```

```{r, echo=F, message=F, warning=F, eval=F}
# Daily format
  # Station / Month / Effort / Covariates / Species      
  tmp <- row.lookup
  tmp$effort <- 1
  # Add species columns  
  tmp[, levels(ind.dat$species)] <- NA
  
  day.obs <- tmp
  day.count <- tmp
# For each week, count the number of individuals/observations
  for(i in 1:nrow(day.obs))
    {
      tmp <- ind.dat[ind.dat$station_id==day.obs$station_id[i] & strftime(ind.dat$date_time, format = "%Y-%m-%d")== day.obs$date[i],]
      
      
      
      for(j in 1:length(levels(ind.dat$species)))
      {
        day.obs[i,levels(ind.dat$species)[j]] <- length(tmp$species[tmp$species==levels(ind.dat$species)[j]])
        day.count[i,levels(ind.dat$species)[j]] <- sum(tmp$event_groupsize[tmp$species==levels(ind.dat$species)[j]])
      }
    }
#write.csv(day.obs, paste0("processed_data/",dat$project_id[1], "_",independent ,"min_Independent_daily_observations.csv"), row.names = F) 

#write.csv(day.count, paste0("processed_data/",dat$project_id[1], "_",independent ,"min_Independent_daily_counts.csv"), row.names = F) 

```
